---
title: "Da Nang, Vietnam — Rooftop Solar PV Potential: Sampling Analysis"
author: "[Your Name]"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
library(dplyr)
library(sf)
library(ggplot2)
library(knitr)
library(kableExtra)
```

## 1. Introduction

This analysis estimates the total rooftop solar PV electricity production
potential for Da Nang, Vietnam using spatial sampling methods.
The methodology follows Grinberg (2022) applied to Rosario, Argentina,
adapted for the Vietnamese urban context.

**Data sources:**
- Building footprints: OpenStreetMap via Geofabrik
- Solar irradiation: Global Solar Atlas v2 (GHI, kWh/m²/yr)
- Land cover strata: Copernicus LULC 2015 (100m)

**Parameters:**
- Panel efficiency (η): 18%
- Performance ratio (PR): 0.75
- Density factor: buildings ≤ 30 m² → elec_prod = 0

---

## 2. Load Data

```{r load-data}
# Load shapefile exported from QGIS/Colab
buildings <- read_sf("../data/danang_solar.shp")

# Reproject to Mercator (consistent with Rosario methodology)
buildings <- st_transform(buildings,
  "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
)

# Apply density factor
buildings$elec_mwh[buildings$area_m2 <= 30] <- 0
buildings$elec_prod[buildings$area_m2 <= 30] <- 0

cat("Total buildings:", nrow(buildings), "\n")
cat("Columns:", paste(names(buildings), collapse=", "), "\n")
```

```{r overview-table}
# Summary table
summary_df <- data.frame(
  Metric  = c("Total buildings", "Buildings with area > 30m²",
              "Total roof area (km²)", "Mean GHI (kWh/m²/yr)",
              "Built-up stratum", "Non built-up stratum"),
  Value   = c(
    format(nrow(buildings), big.mark=","),
    format(sum(buildings$area_m2 > 30), big.mark=","),
    round(sum(buildings$area_m2)/1e6, 2),
    round(mean(buildings$ghi_mean, na.rm=TRUE), 1),
    format(sum(buildings$landuse=="built_up"), big.mark=","),
    format(sum(buildings$landuse=="non_built_up"), big.mark=",")
  )
)
kable(summary_df, caption="Da Nang Building Stock Summary") %>%
  kable_styling(bootstrap_options=c("striped","hover"), full_width=FALSE)
```

---

## 3. Building Area Distribution

```{r area-dist, fig.width=10, fig.height=5}
p1 <- ggplot(buildings %>% filter(area_m2 <= 1000),
             aes(x=area_m2)) +
  geom_histogram(bins=60, fill="#FFAA00", color="white", linewidth=0.2) +
  geom_vline(aes(xintercept=median(area_m2)),
             color="#E8593C", linewidth=1.2, linetype="dashed") +
  annotate("text",
    x=median(buildings$area_m2)+40, y=Inf, vjust=2,
    label=paste0("Median: ", round(median(buildings$area_m2),1), " m²"),
    color="#E8593C", size=3.5) +
  labs(
    title="Building Footprint Area Distribution — Da Nang",
    subtitle=paste0("n = ", format(nrow(buildings),big.mark=","),
                    " buildings (clipped at 1,000 m²)"),
    x="Building Footprint Area (m²)", y="Count"
  ) +
  theme_minimal(base_size=11) +
  theme(plot.title=element_text(face="bold"),
        panel.grid.minor=element_blank())
print(p1)
```

---

## 4. METHOD 1 — Census

```{r census}
N <- nrow(buildings)
total_census <- sum(buildings$elec_mwh, na.rm=TRUE)

cat("=== CENSUS ===\n")
cat("Population N       :", format(N, big.mark=","), "buildings\n")
cat("Total elec (MWh/yr):", format(round(total_census,2), big.mark=","), "\n")
cat("Total elec (GWh/yr):", round(total_census/1000, 2), "\n")
```

---

## 5. METHOD 2 — Random Sampling

```{r sample-size}
# Cochran formula — same as Rosario
Z  <- 1.96   # 95% CI
p  <- 0.5    # max variability
e  <- 0.05   # 5% margin of error
n0 <- (Z^2 * p * (1-p)) / e^2
cat("Cochran formula: n0 =", ceiling(n0), "\n")
n_sample <- 384
cat("Using: n =", n_sample, "\n")
```

```{r random-sampling}
set.seed(42)
sample_rand <- buildings %>% sample_n(n_sample)

# Statistics
y_bar    <- mean(sample_rand$elec_mwh, na.rm=TRUE)
s2       <- var(sample_rand$elec_mwh, na.rm=TRUE)
n        <- nrow(sample_rand)

# Variance of sample mean (finite population correction)
var_ybar <- ((N - n) / N) * (s2 / n)

# Estimated total t̂ = N × ȳ
t_hat    <- N * y_bar
t_var    <- N^2 * var_ybar
set_tot  <- sqrt(t_var)

# 95% CI
CI_rand  <- t_hat + qt(c(0.025, 0.975), df=n-1) * set_tot

cat("=== RANDOM SAMPLING (n=384) ===\n")
cat("Sample mean (ȳ)  :", round(y_bar, 4), "MWh\n")
cat("Estimated total  :", round(t_hat, 1), "MWh/yr\n")
cat("              =  :", round(t_hat/1000, 2), "GWh/yr\n")
cat("95% CI           : [", round(CI_rand[1]/1000,2),
    ",", round(CI_rand[2]/1000,2), "] GWh/yr\n")
```

---

## 6. METHOD 3 — Stratified Sampling

Strata based on Copernicus LULC 2015:
- **Stratum 1 (Built-up):** pixels 50–60 → urban areas
- **Stratum 2 (Non built-up):** all other pixels

```{r strata-overview}
strata_bu    <- buildings %>% filter(landuse == "built_up")
strata_nonbu <- buildings %>% filter(landuse == "non_built_up")

N1 <- nrow(strata_bu)
N2 <- nrow(strata_nonbu)

cat("Stratum 1 — Built-up    :", format(N1, big.mark=","), "buildings\n")
cat("Stratum 2 — Non built-up:", format(N2, big.mark=","), "buildings\n")
```

```{r strata-viz, fig.width=8, fig.height=4}
strata_sum <- buildings %>%
  group_by(landuse) %>%
  summarise(
    n_buildings = n(),
    total_gwh   = sum(elec_mwh, na.rm=TRUE)/1000,
    .groups="drop"
  ) %>%
  mutate(landuse = ifelse(landuse=="built_up",
                          "Built-up\n(pixel 50-60)",
                          "Non built-up\n(other pixels)"))

ggplot(strata_sum, aes(x=landuse, y=total_gwh, fill=landuse)) +
  geom_col(width=0.5, color="#333", linewidth=0.4) +
  geom_text(aes(label=paste0(round(total_gwh,1)," GWh\n(n=",
                              format(n_buildings,big.mark=","),")")),
            vjust=-0.3, size=3.5, fontface="bold") +
  scale_fill_manual(values=c("#FFAA00","#B0D4A0")) +
  labs(title="Solar Potential by Stratum — Da Nang",
       x=NULL, y="Total Electricity Potential (GWh/yr)") +
  theme_minimal(base_size=11) +
  theme(legend.position="none",
        plot.title=element_text(face="bold"),
        panel.grid.major.x=element_blank())
```

### 6a. Equal Size Stratification (60 per stratum)

```{r equal-strat}
calc_stratum <- function(samp, N_s) {
  n_s  <- nrow(samp)
  ybar <- mean(samp$elec_mwh, na.rm=TRUE)
  s2   <- var(samp$elec_mwh, na.rm=TRUE)
  tot  <- N_s * ybar
  varT <- N_s^2 * ((N_s-n_s)/N_s) * (s2/n_s)
  list(n=n_s, mean=ybar, total=tot, var_total=varT)
}

set.seed(42)
s1_eq <- strata_bu    %>% sample_n(min(60, N1))
s2_eq <- strata_nonbu %>% sample_n(min(60, N2))

r1 <- calc_stratum(s1_eq, N1)
r2 <- calc_stratum(s2_eq, N2)

t_eq  <- r1$total + r2$total
se_eq <- sqrt(r1$var_total + r2$var_total)
CI_eq <- t_eq + qt(c(0.025,0.975), df=nrow(s1_eq)+nrow(s2_eq)-2) * se_eq

cat("=== EQUAL STRATIFICATION (60+60=120) ===\n")
cat("Estimated total:", round(t_eq,1), "MWh/yr =",
    round(t_eq/1000,2), "GWh/yr\n")
cat("95% CI: [", round(CI_eq[1]/1000,2),
    ",", round(CI_eq[2]/1000,2), "] GWh/yr\n")
```

### 6b. Optimal Allocation Stratification

```{r optimal-strat}
set.seed(42)
s1_op <- strata_bu    %>% sample_n(min(95, N1))
s2_op <- strata_nonbu %>% sample_n(min(5,  N2))

r1o <- calc_stratum(s1_op, N1)
r2o <- calc_stratum(s2_op, N2)

t_op  <- r1o$total + r2o$total
se_op <- sqrt(r1o$var_total + r2o$var_total)
CI_op <- t_op + qt(c(0.025,0.975), df=nrow(s1_op)+nrow(s2_op)-2) * se_op

cat("=== OPTIMAL STRATIFICATION (95+5=100) ===\n")
cat("Estimated total:", round(t_op,1), "MWh/yr =",
    round(t_op/1000,2), "GWh/yr\n")
cat("95% CI: [", round(CI_op[1]/1000,2),
    ",", round(CI_op[2]/1000,2), "] GWh/yr\n")
```

---

## 7. Comparison of Methods

```{r comparison, fig.width=9, fig.height=5}
results <- data.frame(
  Method = c(
    "Census",
    "Random\n(n=384)",
    "Stratified Equal\n(n=120)",
    "Stratified Optimal\n(n=100)"
  ),
  GWh     = c(total_census/1000,
               t_hat/1000,
               t_eq/1000,
               t_op/1000),
  CI_low  = c(NA, CI_rand[1]/1000, CI_eq[1]/1000, CI_op[1]/1000),
  CI_high = c(NA, CI_rand[2]/1000, CI_eq[2]/1000, CI_op[2]/1000)
)

colors <- c("#1B3A6B","#FFAA00","#E8593C","#2E7D32")

ggplot(results, aes(x=Method, y=GWh, fill=Method)) +
  geom_col(width=0.6, color="#333", linewidth=0.4) +
  geom_errorbar(aes(ymin=CI_low, ymax=CI_high),
                width=0.2, color="#333", linewidth=0.8, na.rm=TRUE) +
  geom_text(aes(label=paste0(round(GWh,1)," GWh")),
            vjust=-0.5, size=3.5, fontface="bold") +
  scale_fill_manual(values=colors) +
  labs(
    title="Rooftop Solar PV Potential — Da Nang, Vietnam",
    subtitle="Comparison of estimation methods (95% CI for sampling methods)",
    x=NULL, y="Electricity Production Potential (GWh/yr)",
    caption="η = 18%  ·  PR = 0.75  ·  GHI = 1,684 kWh/m²/yr"
  ) +
  theme_minimal(base_size=11) +
  theme(legend.position="none",
        plot.title=element_text(face="bold"),
        panel.grid.major.x=element_blank(),
        panel.grid.minor=element_blank())
```

```{r results-table}
results_tbl <- data.frame(
  Method          = results$Method,
  `GWh/yr`        = round(results$GWh, 2),
  `CI Lower (GWh)`= round(results$CI_low, 2),
  `CI Upper (GWh)`= round(results$CI_high, 2)
)
kable(results_tbl, caption="Summary: All Estimation Methods",
      col.names=c("Method","Est. Total (GWh/yr)",
                  "95% CI Lower","95% CI Upper")) %>%
  kable_styling(bootstrap_options=c("striped","hover"), full_width=FALSE) %>%
  row_spec(1, bold=TRUE, background="#e8f0ff")
```

---

## 8. References

- Grinberg, E. (2022). *Estimating building rooftop solar potential in a city
  from a developing country*. MSc Thesis.
- Thompson, S. K. (2012). *Sampling* (3rd ed.). Wiley.
- Global Solar Atlas 2.0, World Bank Group.
- OpenStreetMap contributors, Geofabrik.
- Copernicus Land Monitoring Service, LULC 2015.
